### 2018-05-14 TEX/notes_will/notes_moments/backups/test4_save1.tex

\documentclass{article}
\usepackage{amsthm,amsmath,amssymb}
\usepackage{url} 
\usepackage[authoryear]{natbib}
%%%\usepackage{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{framed}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%%%\usepackage{custom_tex}
\usepackage{grffile}
\usepackage{multirow}
\usepackage{xcolor}
%\usepackage{setspace}
\usepackage{comment}
\usepackage{xr}

\newtheorem{thm}{Theorem}[section]
\newtheorem{definition}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{example}[thm]{Example}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{thm}
\newtheorem{conj}{Conjecture}
\theoremstyle{definition}
\newtheorem{rmk}{Remark}

\newcommand{\F}{\mathcal{F}}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\SUM}{\text{sum}}
\newcommand{\1}{\mathbf{1}}

\newcommand{\mean}[2]{\frac{1}{#1}\sum_{#2=1}^{#1}}

%
%
%
%

\begin{document}
\title{Notes on Big MRA}
\author{William Leeb}
\date{}
\maketitle

\section{Setup}

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = G \ast x + N
%
\end{align}
%
where $G$ is a sum of diracs at locations at least $2L$ apart, and $N$ is a vector of $n$ iid Gaussians. The goal is to recover $x$.

We will assume that the first and last copies of $x$ are at least $L$ places removed from the boundaries of the interval $[1,n + L]$ (if not we can just zero-pad the interval). We will also denote by $I_1,\dots,I_J$ the $J$ subintervals containing the signal; and take $I_j = [a_j,b_j]$. When we take the limit $n \to \infty$, we'll let $J = J_n$ grow with $n$, as is natural. We will see that we require $J_n = \Omega(n)$ in order for the limits to not vanish (and obviously $J_n = O(n)$ too).

For the moments computation, I'm not just computing the expected value. I'm actually showing convergence almost surely as $n \to \infty$. I probably made a mistake in the final formulas since I haven't check them carefully yet; the main point I want to record right now is how to get the bookkeeping correct for dealing with the noise terms, namely, we need to break the averages into averages of averages, each with iid terms.

%
%
%

\section{First moment}

We have:
%
\begin{align}
%
    M_1 = \frac{1}{n} \sum_{i=1}^n Y_i
        = \frac{1}{n/L} \sum_{j=1}^{J_n} \frac{1}{L}\sum_{i=1}^L x_i
            + \frac{1}{n} \sum_{i=1}^n N_i
    \to \gamma \cdot \bar{x},
%
\end{align}
%
where the limit is almost surely (we've used the strong law of large numbers), where $\gamma = \lim_{n\to\infty} J_n L / n$ is the fraction of the observations containing the signal.

\begin{comment}

Actually, what Tamir meant when he said first moment was the average of the windows' averages (a scalar), not the average window (a vector). But this also converges a.s.\ to $\gamma \cdot \bar{x}$, by simply exchanging limits and finite sums, as follows. The sample mean of the window starting at index $i$ is just $\frac{1}{L} \sum_{k=0}^{L-1} Y_{i+k}$. Averaging these over all $n$ windows and taking the limit as $n \to \infty$ we get:
%
\begin{align}
%
    \lim_{n\to\infty} \frac{1}{n} \sum_{i=1}^n \frac{1}{L} \sum_{k=0}^{L-1} Y_{i+k}
    =  \frac{1}{L} \sum_{k=0}^{L-1} \lim_{n\to\infty} \frac{1}{n} \sum_{i=1}^n Y_{i+k}
    = \gamma \cdot \bar{x}.
%
\end{align}

\end{comment}

%
%
%

\section{Second moment}

We fix a value $\Delta$ between $0$ and $L-1$. We will compute the second moment:
%
\begin{align}
%
    M_2(\Delta) = 
        \lim_{n \to \infty} \frac{1}{n} \sum_{i=1}^{n-\Delta} Y_i Y_{i+\Delta}.
%
\end{align}

%
%
%

\subsection{Clean signal without noise}

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
        \to \gamma \cdot R_2(\Delta),
%
\end{align}
%
where here we have defined $R_2(\Delta)$ as element $\Delta$ of the autocorrelation of $x$. 
%
%
%

\subsection{Pure noise without signal}

Here, $Y_i = N_i$. We first fix $\Delta$ between $1$ and $L-1$. Break up the sum into $\Delta$ terms as follows:
%
\begin{align}
%
    M_2(\Delta) &=\frac{1}{n} \sum_{i=1}^n N_i  N_{i+\Delta}
        \nonumber \\
    &= \frac{1}{\Delta} \sum_{m=0}^{\Delta-1} \frac{1}{n/\Delta}\sum_{j=1}^{n/\Delta} 
        N_{j+(j-1)\Delta + m} N_{j + j\Delta + m}.
%
\end{align}
%
Each term
%
\begin{math}
    \frac{1}{n/\Delta}\sum_{j=1}^{n/\Delta} 
        N_{j\Delta + m} N_{(j + 1)\Delta + m}
\end{math}
%
is an average of $n / \Delta$ iid terms with expectation zero, and so converges to 0 a.s.\ as $n \to \infty$. (Of course, it's not exactly $n / \Delta$ terms; there will be finitely many terms unaccounted for this way, but these are negligible).

If $\Delta = 0$, then the computation is even easier:
%
\begin{align}
%
    M_2(\Delta) = \frac{1}{n} \sum_{i=1}^n N_i^2 
        \operatorname*{\longrightarrow}^{a.s.} \sigma^2.
%
\end{align}

So $M_2(\Delta) = \sigma^2$ if $\Delta = 0$, and 0 otherwise.

%%%So the second moment matrix of pure noise converges a.s.\ as $n \to \infty$ to $\sigma^2 I_L$.

%
%
%

\subsection{Signal plus noise}

We will denote by $\XX = x \ast G$, so $Y = \XX + N$. Then the second moment of the signal plus noise is:
%
\begin{align}
%
    M_2(\Delta,Y) = M_2(\Delta,\XX) + M_2(\Delta,N) 
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_{i+\Delta} N_i.
%
\end{align}
%
The law of large numbers says the cross terms vanish as $n \to \infty$. So the limit is simply:
%
\begin{align}
%
    M_2(\Delta,Y) = 
    \begin{cases}
        \gamma \cdot R_2(\Delta), &\text{ if } \Delta > 0; \\
        \gamma \cdot R_2(\Delta) + \sigma^2, &\text{ if } \Delta = 0.
    \end{cases}
%
\end{align}

%
%
%

\section{Third moments}

The same idea lets us compute the third moments. We fix two indices $\Delta_1 < \Delta_2$, and define:
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2) = \sum_{i=1}^{n-\Delta_2} Y_i Y_{i+\Delta_1} Y_{i+\Delta_2}.
%
\end{align}

%
%
%

\subsection{Pure noise, no signal}

Again, the idea is to break up the big average over $n$ terms into a sum of $\Delta_2$ averages, each of $\approx n / \Delta_2$ independent terms. We write:
%
\begin{align}
%
    & M_3(\Delta_1,\Delta_2)  \nonumber \\
    & = \frac{1}{\Delta_2}\sum_{m=0}^{\Delta_2-1}
        \frac{1}{(n/\Delta_2)}\sum_{j=1}^{n/\Delta_2} 
            N_{j+(j-1)\Delta_2+m} N_{j+\Delta_1 + (j-1)\Delta_2 + m} N_{j+j\Delta_2 + m}.
%
\end{align}
%
Each of the $\Delta_2$ terms indexed by $m$ converges is an average of $n/\Delta_2$ independent terms with mean zero, and so converges a.s.\ to zero as $n \to \infty$.

%
%
%

\subsection{Clean signal, no noise}

Again, this is just like the second moment case. Write:
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta_2} 
            Y_i Y_{i+\Delta_1} Y_{i+\Delta_2}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta_2} 
            x_i x_{i+\Delta_1} x_{i+\Delta_2}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta_1} x_{i+\Delta_2}
        \to \gamma \cdot R_3(\Delta_1,\Delta_2),
%
\end{align}
%
where here we have defined $R_3(\Delta_1,\Delta_2)$ as element $(\Delta_1,\Delta_2)$ of the third moment of $x$.

%
%
%

\subsection{Signal plus noise}

We have:
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2,Y) = M_3(\Delta_1,\Delta_2,\XX) + M_3(\Delta_1,\Delta_2,N)
        + \text{cross terms}.
%
\end{align}
%
If $0 < \Delta_1 < \Delta_2$, then all of the cross terms can be shown to go to zero by the same kind of argument we've used several times before.

If $\Delta_1 = \Delta_2 = \Delta > 0$, there is a surviving cross-term, namely:
%
\begin{align}
%
    \frac{1}{n} \sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}^2
    = \frac{1}{(n/L)} \sum_{j=1}^{J_n} \frac{1}{L} \sum_{i=a_j}^{b_j} \XX_i N_{i+\Delta}^2.
%
\end{align}
%
Each of $S_j \equiv \frac{1}{L} \sum_{i=a_j}^{b_j} \XX_i N_{i+\Delta}^2 \sim \frac{1}{L} \sum_{i=1}^{L} x_i \ep_i^2$, where $\ep_i \sim N(0,\sigma^2)$; and they are independent random variables with mean $\E[S_j] = \sigma^2 \cdot \bar{x}$. So by the law of large numbers,
%
\begin{align}
%
    \frac{1}{(n/L)} \sum_{j=1}^{J_n} \frac{1}{L} \sum_{i=a_j}^{b_j} \XX_i N_{i+\Delta}^2
    \to \gamma \cdot \sigma^2 \cdot \bar{x}.
%
\end{align}

Similarly, if $\Delta_2 > \Delta_1 = 0$, then there is a surviving cross-term which converges to $\gamma \cdot \sigma^2 \cdot \bar{x}$ as well. Finally, if $\Delta_1 = \Delta_2 = 0$, then the cross-term converges to $3 \cdot \gamma  \cdot \sigma^2 \cdot \bar{x}$. So in summary, for all $0 \le \Delta_1 \le \Delta_2 \le L-1$, we have:
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2,Y) = 
    \begin{cases}
        \gamma \cdot R_3(\Delta_1,\Delta_2), &\text{ if } 0 < \Delta_1 < \Delta_2; \\
        \gamma \cdot R_3(\Delta_1,\Delta_2) + \gamma \cdot \sigma^2 \cdot \bar{x},
                &\text{ if } \Delta_2 > \Delta_1 = 0 \text{ or } \Delta_1 = \Delta_2 > 0; \\
        \gamma \cdot R_3(0,0) + 3\cdot \gamma \cdot \sigma^2 \cdot \bar{x}
                &\text{ if } \Delta_1 = \Delta_2 = 0.
    \end{cases}
%
\end{align}

%

%
%
%

\section{Attempt at EM}

Let's derive the EM algorithm, where we treat the locations $a_j$ as random latent variables. A simple model is that the $a_j$ are chosen independently and uniformly at random (though take note that this permits overlaps of the signal copies).

We will denote by $G=\sum_{j=1}^J \delta_{a_j}$ the random vector of signal locations. We'll assume the number of signals $J$ is known. The generative model is:
%
\begin{align}
%
    p(Y , G | x)
    = \frac{\exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}}{(2\pi\sigma^2)^{n/2}} 
        \cdot {n \choose J}^{-1}.
%
\end{align}

The log-likelihood is then (up to an additive constant):
%
\begin{align}
%
    \LL(x; Y, G) \propto - \|Y - G\ast x\|^2.
%
\end{align}

The probability distribution of $G$, given $Y$ and $x$, is:
%
\begin{align}
%
    p(G | Y, x) = \frac{ p(Y , G | x) }{ p(Y | x) }
    \propto \exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}.
%
\end{align}

If $x^{(t)}$ is a guess for $x$ on the $t^{th}$ iterate of EM, then we define the weights:
%
\begin{align}
%
    w_t(G) = p(G | Y,x^{(t)}) = \exp\{-\| Y - G \ast x^{(t)} \|^2 / (2\sigma^2)\},
%
\end{align}
%
and the $Q$-function:
%
\begin{align}
%
    Q(x|x^{(t)}) 
    &= \E_{G|Y,x^{(t)}} \left[ \LL(x^{(t)};Y,G) \right]
        \nonumber \\
    &\propto - \sum_{G} \|Y - G\ast x\|^2 \cdot w_t(G).
%
\end{align}
%
The EM algorithm defines the next guess of $x$ to be:
%
\begin{align}
%
    x^{(t+1)} = \operatorname*{\arg\max}_x Q(x|x^{(t)})
        = \operatorname*{\arg\min}_x \sum_{G} \|Y - G\ast x\|^2 \cdot w_t(G).
%
\end{align}

But finding this minimum is hard, since the sum is over all ${n}\choose{J}$ possible values of $G$; if $J \sim n$, as we expect is necessary to have any chance of recovery (this was suggested by the moments method), the number of terms will grow super-exponentially with $n$.

%
%
%

\section{Estimating $\gamma$, using $\sigma$ and the moments}

Using the moments calculation, we will derive an equation relating $\gamma$ and $\sigma$. If $\sigma$ were known (and $\1^\top x \ne 0$), this permits us to then estimate the unscaled moments of $x$, which in turn lets us fit $x$ to its moments.

It will be notationally simpler to instead estimate $\beta \equiv \gamma \cdot L = J / n$. Also, we'll define the unnormalized moments of $x$ by $T_i = L \cdot R_i$, for $i=2,3$. From the first moment we estimate $\beta \cdot (\1^\top x)$, and hence also its square:
%
\begin{align}
%
    A \equiv (\beta \cdot (\1^\top x))^2 
    &= \beta^2  \left( \sum_{i=1}^L x_i^2 + 2 \sum_{i=1}^L \sum_{j=i+1}^{L} x_i x_j \right)
        \nonumber \\
    &= \beta^2 \left( T_2(0) 
            + 2 \sum_{i=1}^{L} \sum_{\Delta=1}^{L-i} x_i x_{i+\Delta}\right)
        \nonumber \\
    &= \beta^2 \left( T_2(0) + 2 \sum_{i=1}^L \sum_{\Delta=1}^{L-1}x_i x_{i+\Delta} 
        \1_{(\Delta \le L-i)} \right)
        \nonumber \\
    &= \beta^2 \left( T_2(0) + 2 \sum_{\Delta=1}^{L-1} \sum_{i=1}^{L-\Delta}x_i x_{i+\Delta} 
        \right)
        \nonumber \\
    &= \beta^2 \left( T_2(0) + 2 \sum_{\Delta=1}^{L-1} T_2(\Delta) \right).
%
\end{align}

On the other hand, from the second moment we can estimate $\beta \cdot T_2(\Delta)$ for all $1 \le \Delta \le L-1$, and $\beta\cdot T_2(0) + \sigma^2$, and consequently $B \equiv \beta \cdot (T_2(0) + 2 \sum_{\Delta=1}^L T_2(\Delta)) + \sigma^2$. Taking the ratio, we obtain the equation:
%
\begin{align}
%
    \frac{1}{\beta} + \frac{\sigma^2}{A} = \frac{B}{A}.
%
\end{align}

If we knew $\sigma$, we can solve for $\beta$:
%
\begin{align}
%
    \beta = \frac{A}{B - \sigma^2}.
%
\end{align}

(Of course, this only works if $B \ne \sigma^2$, or equivalently if $\1^\top x \ne 0$.)

%
%
%
%

\section{A pair of quadratics satisfied by $\beta$}

It turns out that $\beta$ satisfies two independent quadratic equations, defined in terms of the observed third-order moments. The coefficients of these quadratics depend only on the observed moments, not on $\sigma$; so we can estimate $\beta$ (and hence $\gamma$) consistently, even without knowing the noise level. We can then estimate $\sigma$, and from them, recover the moments of $x$, which in turn lets us solve for $x$.

%
%
%
%

\subsection{The first quadratic equation}

The previous section gives the equation:
%
\begin{align}
%
    \beta = \frac{A}{B - \sigma^2}.
%
\end{align}
%

Now we move up to third-order moments. From the first order moments, we can estimate $\beta \cdot \1^\top x$; and from the second-order moments we can estimate $\beta \cdot \|x\|^2 + \sigma^2$. Taking the product, we have:
%
\begin{align}
%
    C &\equiv (\beta \cdot \1^\top x) \cdot (\beta \cdot \|x\|^2 + \sigma^2)
        \nonumber \\
    &= \beta^2 \left( T_3(0,0) + \sum_{\Delta=1}^{L-1}T_3(0,\Delta) 
        + \sum_{\Delta=1}^{L-1}T_3(\Delta,\Delta) \right)
        + \sigma^2 \cdot (\beta \cdot \1^\top x)
        \nonumber \\
    &= \beta^2 \cdot \omega + \sigma^2 \cdot E,
%
\end{align}
%
where 
%
\begin{math}
%
    \omega = T_3(0,0) + \sum_{\Delta=1}^{L-1}T_3(0,\Delta) 
        + \sum_{\Delta=1}^{L-1}T_3(\Delta,\Delta),
%
\end{math}
%
and
%
\begin{align}
%
    E = \beta \cdot \1^\top x.
%
\end{align}

From the observed third-order terms themselves, we can estimate:
%
\begin{align}
%
    D \equiv \beta\cdot \omega + (2L+1)\cdot\sigma^2\cdot (\beta \cdot \1^\top x)
    = \beta\cdot \omega + \sigma^2\cdot F,
%
\end{align}
%
where
%
\begin{align}
%
    F = (2L+1)\cdot(\beta \cdot \1^\top x).
%
\end{align}

Consequently:
%
\begin{align}
%
    D\cdot \beta - \sigma^2 \cdot \beta  \cdot F
    = \beta^2 \cdot w = C - \sigma^2 \cdot E.
%
\end{align}
%
Using $\sigma^2 = B - A/\beta$ and rearranging, we get:
%
\begin{align}
%
    \beta^2 \cdot (D - BF) + \beta \cdot (AF + BE - C) - AE = 0.
%
\end{align}

%
%
%
%

\subsection{The second quadratic equation}

\end{document}

%
%
%
%

\section{Relation to other work on blind deconvolution}

Blind deconvolution is a fundamental and well-studied problem in signal processing. Generally, models of the following form are considered:
%
\begin{align}
%
    Y = h \ast X  + \sigma N,
%
\end{align}
%
where $h$ is a fixed, non-random filter vector of length $L$, say; and $X$ is a length $n$ random vector; and $N$ is a random noise vector. Often, $X$'s entries are assumed to be iid.

Many previous works are focused on recovering the vector $X$ from the observed vector $Y$. A standard method in the $\sigma = 0$ regime involves solving for the unknowns $h$ and $Y$ by maximizing a third or fourth order cumulant of $Y$ \cite{cadzow1996blind}.

\bibliographystyle{plain}
\bibliography{refs_bmra}

\end{document}

\subsection{Second moment of signal plus noise}

Break the sum $\sum_{i=1}^n Y[i + k-1] Y[i + l - 1]$ into three terms: where $i + k-1$ and $i+l-1$ are disjoint from all intervals $I_j$ -- we will denote this set $\I$; where both $i + k - 1$ and $i + l -1$ lie entirely inside one of the $I_j$; and the rest, where exactly one of $i+k-1$ and $i+l-1$ lie in an interval $I_j$. So we have:
%
\begin{align}
%
    & \sum_{i=1}^n Y[i+k-1] Y[i+l-1]    \nonumber \\
    =& \left(\sum_{i \in \I}  
        + \sum_{j=1}^J \sum_{i=a_j - l + 1}^{a_j - k} 
        + \sum_{j=1}^J \sum_{i=b_j - l + 2}^{b_j - k+1} 
        + \sum_{j=1}^J \sum_{i=a_j-k+1}^{b_j-l+1}
      \right)  Y[i+k-1] Y[i+l-1]
%
\end{align}

We will use that $Y[a_j + m] = x[m+1] + \ep[a_j+m]$, $m=0,\dots,L-1$. For the second term, the inner sum is
%
\begin{align}
%
    &\sum_{i=a_j - l + 1}^{a_j - k} Y[i+k-1] Y[i+l-1]
        \nonumber \\
%
    =&\sum_{i=1}^{\Delta} Y[a_j -\Delta + i-1] Y[a_j + i-1]
        \nonumber \\
%
    =& \sum_{i=1}^{\Delta}\ep[a_j -\Delta + i-1] (x[i] + \ep[a_j + i-1])
        \nonumber \\
%
    =& \sum_{i=1}^{\Delta}\ep[a_j -\Delta + i-1] x[i]
        + \sum_{i=1}^{\Delta}\ep[a_j -\Delta + i-1] \ep[a_j + i-1]
%
\end{align}
%

We also use that $Y[b_j - m] = x[L-m] + \ep[b_j-m]$, $m=0,\dots,L-1$. For the third term, the inner sum is
%
\begin{align}
%
    & \sum_{i=b_j - l + 2}^{b_j - k+1} Y[i+k-1] Y[i+l-1]
        \nonumber \\
%
    =& \sum_{i=1}^{\Delta} Y[i+b_j-\Delta] Y[i+b_j]
        \nonumber \\
%
    =& \sum_{i=1}^{\Delta} (x[L-\Delta+i]+\ep[i+b_j-\Delta]) \ep[i+b_j]
        \nonumber \\
%
    =& \sum_{i=1}^{\Delta} x[L-\Delta+i] \ep[i+b_j]
        + \sum_{i=1}^{\Delta} \ep[i+b_j-\Delta] \ep[i+b_j]
\end{align}
%

The inner sum of the fourth term is:
%
\begin{align}
%
    &\sum_{i=a_j-k+1}^{b_j-l+1} Y[i+k-1] Y[i+l-1]
        \nonumber \\
%
    =& \sum_{i=1}^{L - \Delta} Y[i+a_j-1] Y[i+a_j + \Delta - 1]
        \nonumber \\
%
    =& \sum_{i=1}^{L - \Delta} (x[i]+\ep[i+a_j-1])(x[i+\Delta] +\ep[i+a_j + \Delta - 1])
        \nonumber \\
%
    =& \sum_{i=1}^{L - \Delta} x[i] x[i+\Delta]
        + \sum_{i=1}^{L - \Delta} x[i]\ep[i+a_j + \Delta - 1]
        \nonumber \\
        &+ \sum_{i=1}^{L - \Delta} \ep[i+a_j-1] x[i+\Delta] 
        + \sum_{i=1}^{L - \Delta} \ep[i+a_j-1]\ep[i+a_j + \Delta - 1])
%
\end{align}

\end{document}